---
title: "Graph Display"
author: "Perrine Kergoat"
date: "2023-12-01"
output:
  pdf_document: 
    latex_engine: xelatex
---

```{r message=FALSE, warning=FALSE, include=FALSE, paged.print=TRUE}
knitr::opts_chunk$set(echo = TRUE)

library(vcfR)
library(hierfstat)
library(gaston)

load("/Users/perrinekergoat/Desktop/CheckBeta/Environments/Variance/Graph_testFST_500_35739448.RData")
load("~/Desktop/CheckBeta/Environments/Other_methods/FST_calc_50035739448.Rdata") 


#' Calculate the mean of each group of n elements in a vector,
#' with an optional step size
#'
#' @param vecteur a numeric vector
#' @param n a positive odd integer indicating the number of elements to consider
#' for each group. The number has to be odd for computing reasons
#' (could be improved in the future)
#' @param slide an integer indicating the number of elements to slide the window
#' forward at each iteration (default 1)
#'
#' @return a numeric vector containing the means of each group of n elements
#'
#' @examples
#' mean_groups(1:10, 3, 1)
#'
#' @export
SlidingMean <- function(vecteur, n, slide=1) {
  # Check if number is odd
  if((n %% 2) == 0) {
    print(paste(n,"is Even, please provide an odd number"))
    break
  }
  
  if(n > 1) {
    WinSizeSide = ((n-1)/2)
  }
  else {
    WinSizeSide = n
  }
  
  # Create a vector full of 0 for the means
  means <- rep(0,length(vecteur))
  
  # Loop through all groups of n elements and calculate the mean
  for (i in seq(from=WinSizeSide+1, to=length(vecteur)-n+WinSizeSide, by=slide)) {
    means[i] <- mean(vecteur[(i-WinSizeSide):(i+WinSizeSide)])
  }
  
  # Return the vector of mean
  return(means)
}

#position <- (match(5e6, vcf_file@snps$pos))


tot_min <- min(c(min(pop_summary$"mean Beta"), min(pop_summary$"max Beta"), min(pop_summary$"min Beta")))
tot_max <- max(c(max(pop_summary$"mean Beta"), max(pop_summary$"max Beta"), max(pop_summary$"min Beta")))
```

# Parameters

-   id : 35739463
-   chromosome size : 1e7 - 1
-   number of generations : 900
-   dominance : 0.8
-   selection : 0.030
-   migration : 0.001
-   recombination : 1e-8
-   *Allele* frequency in subpopulation 1 : `r freq_check[1]`
-   *Allele* frequency in subpopulation 2 : `r freq_check[2]`
-   global Fst : `r Fst_calc$Fst`

# Fst

## Whole genome

```{r}
#position <- (match(5e6, vcf_file@snps$pos))
position <- 5e6
meanFST_500 = SlidingMean(vecteur = Fst_calc$FstSNP,n = 5001, slide = 101)
str(Fst_calc)

plot(Fst_calc$FstSNP ~ vcf_file@snps$pos, pch = 19, xlab = "Genomic position", ylab = "Pairrwise Fst")
abline(v = position, col = "red")
points(position,Fst_calc$FstSNP[match(5e6, vcf_file@snps$pos)], col = "red", cex = 2)
#lines(meanFST_500, col = "cyan")

#legend("topright", legend = c("Mutation position", "Fst mean (wind. of 501 SNPs)"), fill = c("red", "cyan"))

```

## Filtered whole genome

```{r}
#meanFST_500 = SlidingMean(vecteur = Fst_calc_maf_01$FstSNP,n = 501, slide = 1)
#meanFST_200 = SlidingMean(vecteur = Fst_calc_maf_01$FstSNP,n = 201, slide = 1)
#plot(Fst_calc_maf_01$FstSNP, pch = 19)
#lines(meanFST_500, col = "cyan")
#lines(meanFST_200, col = "green")
#legend("topright", legend = c("Mutation position", "Fst mean (wind. of 501 SNPs)"), fill = c("cyan", "green"))

```

## Zoom around mutation

```{r}
#plot(Fst_calc$FstSNP, pch = 19, xlim = c(position-5000, position+5000) ) #, main = (paste0("Dominance : 0.8, sel coeff: 0.025, chromosome : 1e7-1, recomb : 1e-8, migration : 0.001, \n freq in the 2 pop :", freq_check[1], " and ", freq_check[2], ", Fst global ", Fst_calc$Fst)))
#abline(v = position, col = "red")
#points(position,Fst_calc$FstSNP[position], col = "red", cex = 3)
#lines(meanFST_500, col = "cyan", lwd = 4)
#lines(meanFST_200, col = "green", lwd = 4)
#legend("topright", legend = c("Mutation position", "Fst mean (wind. of 501 SNPs"), fill = c("red", "cyan"))

```

# Pi

## Whole population

```{r}
#plot(pop_summary$"pi", type = "l", col = "firebrick4")
#abline(v = list_start[[1]], col = "black")
#abline(v = tail(list_start, n = 1), col = "black")

```

## 2 subpopulations separated

```{r}
genomic_position <- list()
i <- 0 
for (ele in start_bound) {
  i <- i + 1 
  genomic_position[i] <- vcf_file@snps$pos[ele]
}

plot(pop1_summary$pi ~ unlist(genomic_position), col = "purple", type = "l", ylim = c(min(pop2_summary$pi - pop1_summary$pi - 10), max(pop1_summary$pi) + 10), xlab = "Genomic position", ylab = "pi")
lines(pop2_summary$pi ~ unlist(genomic_position), col = "cyan")
abline(v = genomic_position[[list_start[[1]]]], col = "grey")
abline(v = genomic_position[[list_start[[4]]]], col = "grey")
legend("bottomright", legend = c("pop1", "pop2", "delta-pi"), fill = c("purple", "cyan", "darkorange"))
lines(pop2_summary$pi - pop1_summary$pi ~ unlist(genomic_position), col = "darkorange")

```

# Simple metrics of Beta Matrix

## Variance

```{r}
plot(pop_summary$`variance Beta` ~ unlist(genomic_position), type = "l", col = "chartreuse3", xlab = "Genomic position", ylab = "Beta variance")
abline(v = genomic_position[[list_start[[1]]]], col = "grey")
abline(v = genomic_position[[list_start[[4]]]], col = "grey")

```

# Top 5% values

```{r}
#snps_fst <- as.vector(na.omit(Fst_calc$FstSNP))
  ### Fst filtered by window
#j <- 0
#snps_window <- list()

#for (bound in start_bound){
  #j <- j + 1 
  #snps_window[[j]] <- mean(snps_fst[as.integer(bound) : as.integer(bound + 99)])
#}

#data_fst <- as.numeric(snps_fst)
  ###
#df_fst <- nb_windows
#perc_fst95 <- qchisq(p = 0.95, df = df_fst)
#perc_fst5 <- qchisq(p = 0.05, df = df_fst)
#out_fst95 <- data_fst[data_fst > perc_fst95]
#out_fst5 <- data_fst[data_fst < perc_fst5]

#data_diff_pi <- pop2_summary$pi - pop1_summary$pi
#qqPlot(data_diff_pi)
#qqline(data_diff_pi)

#data_pi <- pop_summary$pi
#qqPlot(data_pi)
#qqline(data_pi)
#mean_pi <- mean(data_pi)
#sd_pi <- sd(data_pi)
#perc_pi95 <- qnorm(0.99, mean = mean_pi, sd = sd_pi)
#lim_pi95 <- pnorm(perc_pi95, mean = mean_pi, sd = sd_pi)
#perc_pi5 <- qnorm(0.05, mean = mean_pi, sd = sd_pi)
#out_pi95 <- data_pi[data_pi > perc_pi95]
#out_pi5 <- data_pi[data_pi < perc_pi5]

#plot(data_pi)
#abline(h = perc_pi95, col = "red")

#data_beta <- pop_summary$`variance Beta`
#df_beta <- nb_windows
#perc_beta95 <- qchisq(p = 0.95, df = df_beta)
#lim_beta95 <- pchisq(perc_beta95, df = df_beta)
#perc_beta5 <- qchisq(p = 0.05, df = df_beta)
#lim_beta5 <- pchisq(perc_beta5, df = df_beta)
#out_beta95 <- data_beta[data_beta > perc_beta95]
#out_beta5 <- data_beta[data_beta < perc_beta5]

#plot(data_beta)
#lines(x = perc_beta95)

#plot(data_pi)
#abline(h = perc_pi95, col = "red")
#plot(density(data_pi))
#abline(v = perc_pi95, col = "red")

#plot(data_beta)
#abline(h = perc_beta95, col = "red")
#plot(density(data_beta))
#abline(v = perc_beta95, col = "red")

#plot(density(rchisq(1000, nb_windows)))
#abline(v = perc_beta95, col = "red")


#plot(density(data_fst))
#abline(v = perc_fst95, col = "red")
#plot(density(data_beta))
#abline(v = perc_beta95, col = "red")
#plot(density(rchisq(10000, nb_windows)))
#abline(v = perc_beta95, col = "red")

#plot(data_fst)
#plot(data_beta)

#ks.test(data_fst, "pnorm")
```

# Top 5% values

```{r}
percent <- 1

### Fst
snps_fst <- as.vector(na.omit(Fst_calc$FstSNP))

  ### Fst filtered by window
j <- 0
snps_window <- list()

for (bound in start_bound){
  j <- j + 1 
  snps_window[[j]] <- mean(snps_fst[as.integer(bound) : as.integer(bound + 99)])
}
  ###

data_fst <- as.numeric(snps_window)

top5_fst <- tail(sort(data_fst, decreasing = FALSE), n = round(length(data_fst) * percent / 100), digits = 0)
lim_fst95 <- min(top5_fst)

plot(data_fst ~ unlist(genomic_position), main = "Fst", xlab = "Genomic position", ylab = "Fst")
abline(h = lim_fst95, col = "deepskyblue")
abline(v = 5e6 -1)
legend("topright", legend = c("Top 5 %"), fill = c("deepskyblue"))
plot(density(data_fst), main = "Fst density")
abline(v = lim_fst95, col = "deepskyblue")
legend("topright", legend = c("Top 5 %"), fill = c("deepskyblue"))


### Pi

data_pi <- pop_summary$pi

top5_pi <- tail(sort(data_pi, decreasing = TRUE), n = round(length(data_pi) * percent / 100), digits = 0)
lim_pi95 <- min(top5_pi)

  ### Genomic positions
genomic_position <- list()
i <- 0 
for (ele in start_bound) {
  i <- i + 1 
  genomic_position[i] <- vcf_file@snps$pos[ele]
}
  ###

plot(data_pi ~ unlist(genomic_position), main = "pi", xlab = "Genomic position", ylab = "pi")
abline(h = lim_pi95, col = "purple")
abline(v = 5e6 -1)
legend("topright", legend = c("Top 5 %"), fill = c("purple"))

plot(density(data_pi), main = "pi density")
abline(v = lim_pi95, col = "purple")
legend("topright", legend = c("Top 5 %"), fill = c("purple"))

### Beta variance

data_beta <- pop_summary$`variance Beta`

top5_beta <- tail(sort(data_beta, decreasing = FALSE), n = round(length(data_pi) * percent / 100), digits = 0)
lim_beta95 <- min(top5_beta)

plot(data_beta ~ unlist(genomic_position), main = "Beta variance", xlab = "Genomic position", ylab = "Beta variance")
abline(h = lim_beta95, col = "deeppink")
abline(v = 5e6 -1)
legend("topright", legend = c("Top 5 %"), fill = c("deeppink"))
plot(density(data_beta), main = "Beta variance density")
abline(v = lim_beta95, col = "deeppink")
legend("topright", legend = c("Top 5 %"), fill = c("deeppink"))
```
