---
title: "Master_project"
author: "Perrine Kergoat"
date: "2023-09-28"
output:
  pdf_document: 
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Open VCF file
```{r}
library(vcfR)
library(hierfstat)
library(gaston)
library(stats)
library(StatMatch)
library(dgof)
library(outliers)
library(nortest)

vcf_file <- read.VCF("/Users/perrinekergoat/Master_Project/Code/Data/Simu.vcf")
print(vcf_file)



```

```{r}
#####
plot(vcf_file@snps$pos,vcf_file@snps$pos)
abline(h=5000000, col ="red")

which(vcf_file@snps$pos==500000)
# 3351

i = 0
for (start in start_bound){
  i <- i + 1
  end <- start + l_window
  
  if (end > nb_snps) {end = nb_snps}
  if (start == nb_snps) {break}
  
  abline(h = min(vcf_file@snps$pos[start:end]))
  abline(h = max(vcf_file@snps$pos[start:end]))
}

for (i in 0:20){
S = 2000
start = S + (100*i)
end = start + 1000
reduced_vcf <- vcf_mat[,start:end]
kinship_calc <- beta.dosage(reduced_vcf)
image(kinship_calc[c(1:200)*10,c(1:200)*10], main = paste0(start,":",end))
}


###

M = matrix(c(1:100), ncol = 10)
M[upper.tri(M, diag = TRUE)]

#####
```



### Convert VCF file to matrix
```{r}
vcf_mat <- as.matrix(vcf_file)
beta_mat <- beta.dosage(vcf_mat)
``` 

### Get SNPs and parameters to cut the genomes
```{r}
data_snps <- vcf_file@snps
nb_snps <- nrow(data_snps)
nb_windows <- nb_snps %/% 1000 # nb of snps in windows
l_window <- nb_snps %/% nb_windows 
start_bound <- seq(from = 1, to = nb_snps, by = l_window)
#end_bound[6] <- nb_snps

print(nb_snps)
print(nb_windows)
print(l_window)

print(start_bound)

#reduced_vcf <- vcf_mat[,start:end]
#kinship_calc <- beta.dosage(reduced_vcf)




```

### Create windows in the dosage matrix 
```{r}
list_beta <- list()
i <- 0 

par(mfrow = c(2,3))
for (start in start_bound){
  i <- i + 1
  print(i)
  end <- start + l_window
  
  if (end > nb_snps) {end = nb_snps}
  if (start == nb_snps) {break}
  if (end-start <= nb_windows) {break}
  else {
    reduced_vcf <- vcf_mat[,start:end]
    kinship_calc <- beta.dosage(reduced_vcf)
    #print(typeof (test))
    list_beta[[i]] <- kinship_calc
    image(kinship_calc[c(1:200)*10,c(1:200)*10], main = paste0("Window",":",i, " (start at: ", start, " end at: ", end))
  }
  
  
}

#print(test)
```

### Windows matrix output
```{r}
par(mfrow = c(2,3))
for (num in 1:nb_windows) {
  name_file <- paste0("/Users/perrinekergoat/Master_Project/Code/Data/HM-window nb ",num,".png")
  png(file = name_file, width = 2000, height = 2000)
  heatmap(list_beta[[num]], Colv = NA, Rowv = NA, scale = "none", keep.dendro = FALSE, main = paste0("Window",":",num)) # enlever le tri des individus 
  dev.off()
}

#image(kinship_calc[c(1:200)*10,c(1:200)*10], main = paste0(start,":",end))
```

### Whole genome matrix output 
```{r}
png(file="/Users/perrinekergoat/Master_Project/Code/Data/HM-global.png", width=20000, height=20000)
heatmap(beta_mat, Colv = NA, Rowv = NA, scale = "none", keep.dendro = FALSE) # enlever le tri des individus 
#dev.off()

image(beta_mat[c(1:200)*10,c(1:200)*10])
```

### Mahalanobis distance
```{r}
#mahalanobis.dist(data.x = list_dos[[1]], data.y = list_dos[[2]])

#mahala_results <- mahalanobis.dist(data.x = list_dos[[1]], data.y = list_dos[[2]])
```

### Convert beta matrices in vectors (onnly half of the matrices)
```{r}
list_half_beta <- list()

for (j in 1:nb_windows){
  #print(j)
  list_half_beta[[j]] <- list_beta[[j]][upper.tri(list_beta[[j]], diag = TRUE)]
}

list_half_gen <- upper.tri(beta_mat, diag = TRUE)

```




### Cut the genomes in windows 
```{r}
list_diff1 <- list()
list_diff2 <- list()
list_mean1 <-list()
list_mean2 <- list()
tot_diff1 <- list()
tot_diff2 <- list()

for (k in 1:nb_windows){
  m <- 0
  list_diff <- list()
  for (l in 1:nb_windows){
    if (l != k){
      diff_beta1 <- abs(list_half_beta[[k]] - list_half_beta[[l]])
      diff_beta2 <- sqrt(list_half_beta[[k]]^2 + list_half_beta[[l]]^2)
      #diff_beta3 <- mahalanobis.dist(data.x = list_half_beta[[k]], data.y = list_half_beta[[l]])
      
      list_mean1[[k]] <- mean(diff_beta1)
      list_mean2[[k]] <- mean(diff_beta2)
      
      list_diff1[[k]] <- diff_beta1
      list_diff2[[k]] <- diff_beta2
      #print(diff_beta[1:50])
      
      m <- m + 1
    }
    #print(paste0("list ", k, " minus list ", l))
    #print(length(list_diff))
    
  }
  tot_diff1[[k]] <- Reduce("+", list_diff1)
  tot_diff2[[k]] <- Reduce("+", list_diff2)
  print(k)
}

tot_sum1 <- Reduce("+", tot_diff1)
tot_sum2 <- Reduce("+", tot_diff2)
```


```{r}
diff_win1 <- list()
diff_win2 <- list()

for (k in 1:nb_windows){
  
  diff_win1[k] <- abs(list_half_beta[[k]] - list_half_gen)
  diff_win2[k] <- sqrt(list_half_beta[[k]]^2 + list_half_gen^2)
}

```


### plot the distribution of the window and the mean one 
```{r}
add.alpha <- function(col, alpha=1){
  if(missing(col))
    stop("Please provide a vector of colours.")
  apply(sapply(col, col2rgb)/255, 2, 
        function(x) 
          rgb(x[1], x[2], x[3], alpha=alpha))  
}


#data1 <- c(tot_diff[[4]])
#data0 <- c(tot_sum)
#hist(x = data1, col = add.alpha("cyan",.8), breaks = c(0:75), ylim = c(0,1700000))
#hist(x = data2, col = add.alpha("purple",.5), add = TRUE, breaks = c(0:75))

data_glob1 <- Reduce("+", tot_diff1)/nb_windows
data_glob2 <- Reduce("+", tot_diff2)/nb_windows

par(mfrow = c(2,3))
for (i in 1:nb_windows){
  S = 1
  start = S + (10*i)
  end = start + 1000
  data <- paste0("data", i)
  hist(x = tot_diff1[[i]], col = "cyan", main = paste0("Window:",i, " (addition)"), xlim = c(0,200), ylim = c(0,1e6))
  hist(x = data_glob1, col = add.alpha("purple",.2), add = TRUE)
}
hist(x = data_glob1, col = add.alpha("purple",.2))

par(mfrow = c(2,3))
for (i in 1:nb_windows){
  S = 1
  start = S + (10*i)
  end = start + 1000
  data <- paste0("data", i)
  hist(x = tot_diff2[[i]], col = "cyan", main = paste0("Window:",i, " (sqrt)"), xlim = c(0,200), ylim = c(0,1e6))
  hist(x = data_glob2, col = add.alpha("purple",.2), add = TRUE)
}
hist(x = data_glob2, col = add.alpha("purple",.2))








par(mfrow = c(2,3))
for (i in 1:nb_windows){
  S = 1
  start = S + (10*i)
  end = start + 1000
  data <- paste0("data", i)
  hist(x = diff_win1[[i]], col = "cyan", main = paste0("Window:",i, " (addition)"), xlim = c(0,200), ylim = c(0,1e6))
  hist(x = list_half_gen, col = add.alpha("purple",.2), add = TRUE)
}
hist(x = data_glob1, col = add.alpha("purple",.2))

par(mfrow = c(2,3))
for (i in 1:nb_windows){
  S = 1
  start = S + (10*i)
  end = start + 1000
  data <- paste0("data", i)
  hist(x = diff_win2[[i]], col = "cyan", main = paste0("Window:",i, " (sqrt)"), xlim = c(0,200), ylim = c(0,1e6))
  hist(x = data_glob2, col = add.alpha("purple",.2), add = TRUE)
}
hist(x = list_half_gen, col = add.alpha("purple",.2))


```


### Kolmogorov-Smirnov Test
```{r}
results_ks1 <- list()
tot_resks1 <- list()
results_ks2 <- list()

for (n in 1:nb_windows) {
  res <- ks.test(x = data_glob1, y = tot_diff1[[n]])
  tot_resks1[[n]] <- res
  results_ks1[[n]] <- res$statistic[[1]]
  print(n)
}

for (n in 1:nb_windows) {
  res <- ks.test(x = data_glob2, y = tot_diff2[[n]])
  results_ks2[[n]] <- res$statistic[[1]]
  print(n)
}

```

```{r}
tot_resks1[[1]]
```


```{r}
plot(as.numeric(results_ks1))
plot(as.numeric(results_ks2))

hist(as.numeric(results_ks1))
hist(as.numeric(results_ks2))

typeof(results_ks1)
```


### Detect if a window is significantly further from the mean of all other windows
```{r}
grubb_res1 <- grubbs.test(as.numeric(results_ks1), type = 10)
if (grubb_res1$p.value < 0.05){
  w_interest1 <- which.max(results_ks1)
  print(w_interest1)
}

max_w1 <- which.max(results_ks1)
print(paste0("Not significant (addition): ", max_w1))


grubb_res2 <- grubbs.test(as.numeric(results_ks2), type = 10)
if (grubb_res2$p.value < 0.05){
  w_interest2 <- which.max(results_ks2)
  print(w_interest2)
}

max_w2 <- which.max(results_ks2)
print(paste0("Not significant (sqrt): ", max_w2))
```


### Using mean of the disatance of each window
```{r}
par(mfrow = c(1,2))
plot(as.numeric(list_mean1), main = "Addition")
plot(as.numeric(list_mean2), main = "Sqrt")

par(mfrow = c(1,2))
hist(as.numeric(list_mean1), main = "Addition")
hist(as.numeric(list_mean2), main = "Sqrt")

grubb_mean1 <- grubbs.test(as.numeric(list_mean1), type = 10)
if (grubb_mean1$p.value < 0.05){
  w_interest_m1 <- which.max(list_mean1)
  print(w_interest_m1)
}

grubb_mean2 <- grubbs.test(as.numeric(list_mean2), type = 10)
if (grubb_mean2$p.value < 0.05){
  w_interest_m2 <- which.max(list_mean2)
  print(w_interest_m2)
}


```























### Get the corresponding place to look at in the genome 
```{r}

```




### Essais
```{r}
norm_test1 <- ad.test(data_glob1)
norm_test1

norm_test2 <- ad.test(data_glob2)
norm_test2
```







```{r}
vcf_file@snps$id[vcf_file@snps$pos == 5e6]

```


### Brouillon  
```{r}
a <- c(2,3,4,5)
b <- c(1,1,1,1)

c <- a-b
print(c)

print(a^2)
print(sqrt(4*b))

M = matrix(c(1:100), ncol = 10)
M[upper.tri(M, diag = TRUE)]

```

- comparisons to the global matrice 
- 





